---
title: "Reproducible Research: Course Project 1"
Author:
        "Omer Shechter"
Date:
        "20/09/2018"
output: 
  html_document:
    keep_md: true
---
## Data Loading
Download the files and unzip them into a data directory 
Check if data directory exist, if not create it
Than download the data and unzip it if not allredy exist
Read the Dataset into a datframe - Activity_DF
```{r Load Data}
if(!file.exists("data")) {  
        dir.create("data")  
}  
if(!file.exists("./data/Activity.zip")) {  
        fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"  
        download.file(fileUrl, destfile="./data/Activity.zip")  
        if(file.exists("./data/Activity.zip"))   
                unzip("./data/Activity.zip",exdir="./data")  
}
Activity_DF<-read.csv("./data/Activity.csv",sep = ",")
```

## First look at the dataset

```{r Data Summary}
str(Activity_DF)
#Show only the non missing values in steps 
head(Activity_DF[which(!is.na(Activity_DF$steps)), ])
summary(Activity_DF)
```


## What is mean total number of steps taken per day?


Group the number of steps by date and intervals. 
Find the total number of steps per day over all days. 
some of the days have no steps data. 
This will be not shwon in the plot.

```{r}
#Create a dataframe with  total steps per day  and dates
Steps_Per_Day<-data.frame(date = unique(as.Date(Activity_DF$date,"%Y-%m-%d")), Total.Steps = tapply(Activity_DF$steps,as.Date(Activity_DF$date,"%Y-%m-%d"),sum))

head(Steps_Per_Day)
```

Plot the Histogram of the total steps per day (Including NA values)
```{r Histogram Plot}
##Load ggplot2
library(ggplot2)

#Plot the Histogram (with missing values)
qplot(Steps_Per_Day$Total.Steps, geom="histogram",bins=10,xlab = "Steps",ylab = "Count",main = "Total Steps per day Histogram")  

```



## What is the average daily activity pattern?

Calculate the mean and median per day 
```{r Process}
#Add the mean and median steps per day to the Steps_Per_Day
Steps_Per_Day<-cbind(Steps_Per_Day,data.frame(Steps.mean = tapply(Activity_DF$steps,as.Date(Activity_DF$date,"%Y-%m-%d"),mean)))

Steps_Per_Day<-cbind(Steps_Per_Day,data.frame(Steps.median = tapply(Activity_DF$steps,as.Date(Activity_DF$date,"%Y-%m-%d"),median)))
head(Steps_Per_Day)

```




## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
